The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-03-01 src/ms.tex
improvements to Astropy Fit + Time
Co-authored-by: Aarya Patil <aaryapatil1996@gmail.com>
Signed-off-by: nstarman <nstarkman@protonmail.com>

\newcommand{\astropySpectralCoord}{\astropyapidoc{coordinates.SpectralCoord}{SpectralCoord}}
\newcommand{\astropySkyCoord}{\astropyapidoc{coordinates.SkyCoord}{SkyCoord}}
\newcommand{\astropyICRS}{\astropyapidoc{coordinates.builtin_frames.ICRS}{ICRS}}
\newcommand{\astropyGalacticLSR}{\astropyapidoc{coordinates.builtin_frames.
GalacticLSR}{GalacticLSR}}
\newcommand{\astropyAltAz}{\astropyapidoc{coordinates.builtin_frames.AltAz}{AltAz}}
\newcommand{\astropyGalactocentric}{\astropyapidoc{coordinates.builtin_frames.Galactocentric}{Galactocentric}}
\newcommand{\astropyCosmology}{\astropyapidoc{cosmology.Cosmology}{Cosmology}}
\newcommand{\astropyFlatLambdaCDM}{\astropyapidoc{cosmology.FlatLambdaCDM}{FlatLambdaCDM}}
\newcommand{\astropyTable}{\astropyapidoc{table.Table}{Table}}
\newcommand{\astropyQTable}{\astropyapidoc{table.QTable}{QTable}}
\newcommand{\astropyTime}{\astropyapidoc{time.Time}{Time}}
\newcommand{\astropyLeapSeconds}{\astropyapidoc{time.LeapSeconds}{LeapSeconds}}
\newcommand{\astropyDistribution}{\astropyapidoc{uncertainty.Distribution}{Distribution}}
\newcommand{\astropyUnit}{\astropyapidoc{units.Unit}{Unit}}
\newcommand{\astropyQuantity}{\astropyapidoc{units.Quantity}{Quantity}}
\newcommand{\astropyScienceState}{\astropyapidoc{utils.state.ScienceState}{ScienceState}}

@@

Summarize major pieces of development since the last paper (since v2.0). A few
ideas (taken from What's New pages) below, but feel free to expand on or remove
from this list:
\begin{itemize}
    \item Support for representing and transforming velocity data in coordinates, epoch propagation (v3.0)
    \item Improved support for astronomical time series: TimeSeries object (v3.2), Box Least Squares periodogram (v3.1)
    \item Overall improved support of Quantity throughout numpy (v4.0) and scipy
    \item Native support for Time, Quantity, and SkyCoord objects in Astropy tables
    \item APE 14 WCS
    \item A new SpectralCoord class for representing and transforming spectral quantities
    \item Support for dask/large arrays in several places in the core package (4.1, 5.0, and maybe other versions)
\end{itemize}

>>>>

\newcommand{\astropySpectralCoord}{\astropyapidoc{coordinates.SpectralCoord}{SpectralCoord}}
\newcommand{\astropySkyCoord}{\astropyapidoc{coordinates.SkyCoord}{SkyCoord}}
\newcommand{\astropyICRS}{\astropyapidoc{coordinates.builtin_frames.ICRS}{ICRS}}
\newcommand{\astropyGalacticLSR}{\astropyapidoc{coordinates.builtin_frames.
GalacticLSR}{GalacticLSR}}
\newcommand{\astropyAltAz}{\astropyapidoc{coordinates.builtin_frames.AltAz}{AltAz}}
\newcommand{\astropyGalactocentric}{\astropyapidoc{coordinates.builtin_frames.Galactocentric}{Galactocentric}}
\newcommand{\astropyCosmology}{\astropyapidoc{cosmology.Cosmology}{Cosmology}}
\newcommand{\astropyFlatLambdaCDM}{\astropyapidoc{cosmology.FlatLambdaCDM}{FlatLambdaCDM}}
\newcommand{\astropyFitsColumn}{\astropyapidoc{io.fits.Column}{Column}}
\newcommand{\astropyModel}{\astropyapidoc{modeling.Model}{Model}}
\newcommand{\astropyTableColumn}{\astropyapidoc{table.Column}{Column}}
\newcommand{\astropyTable}{\astropyapidoc{table.Table}{Table}}
\newcommand{\astropyQTable}{\astropyapidoc{table.QTable}{QTable}}
\newcommand{\astropyTime}{\astropyapidoc{time.Time}{Time}}
\newcommand{\astropyLeapSeconds}{\astropyapidoc{time.LeapSeconds}{LeapSeconds}}
\newcommand{\astropyDistribution}{\astropyapidoc{uncertainty.Distribution}{Distribution}}
\newcommand{\astropyUnit}{\astropyapidoc{units.Unit}{Unit}}
\newcommand{\astropyQuantity}{\astropyapidoc{units.Quantity}{Quantity}}
\newcommand{\astropyMasked}{\astropyapidoc{utils.masked.Masked}{Masked}}
\newcommand{\astropyScienceState}{\astropyapidoc{utils.state.ScienceState}{ScienceState}}

@@

  The \astropypkg core package is mature and stable, in that many features have
  been part of the core package for years with (purposefully) few changes to the
  software interface. This maturity and stability allows for the broader
  astronomy \python community to rely upon the \astropypkg core package and
  build specialized packages within the Astropy ecosystem. Even many
  non-astronomical \python libraries have come to rely on \astropypkg.

  The relationship between \astropy and community is reciprocal, with the
  growing and evolving needs of the community and the maturation of the broader
  scientific \python ecosystem driving much of the development of the
  \astropypkg core package.

  One manifestation of this was the realization that some features do not
  actually belong in the core package, either because they are of more general
  use outside of \astropypkg, or because they are too specialized. An example of
  the former is the copy of the IAU Standards Of Fundamental Astronomy (SOFA)
  software Collection\footnote{\url{http://www.iausofa.org}} \citep{sofa} that
  \astropy carried, as well as the \python wrappers for it.  Since this is more
  basic infrastructure, with a release cycle set by SOFA, it made more sense to
  create separate packages that \astropypkg can depend on: \package{ERFA}
  \citep{erfa} and \package{PyERFA} \citep{pyerfa}. An example of something that
  was too specialized was a Virtual Observatory sub-package (now removed).  Of
  its content, the Simple Application Message Protocol made sense to keep
  (\astropysubpkg{samp}), but other parts belonged more properly in \package{
  astroquery} (see Sect.~\ref{sec:astroquery}).

  Within \astropypkg, development may be roughly split into a few categories:
  new features added, intra and inter-package interoperability, improvements to
  precision, accuracy, and reproducibility, and performance enhancements. We
  discuss these in turn.

\subsubsection*{New and Planned Features} \label{sec:core-features-new}

New features come in various forms, from new submodules providing
wholly new capabilities, to larger additions to existing submodules,
to large rewrites of the interface of modules.  This range is spanned
by the five main new features that entered the \astropypkg core since the
previous summary \citepalias{astropy:2018}.

New features present a particular difficulty for a mature package such
as \astropypkg, because by their nature they have not been tried out
much and it is difficult to know in advance what the best interface
will be, and hard to address all possible use cases in one go, even
within \astropypkg\ itself.  To give a sense of this, we include in
our descriptions how we hope the new features will evolve (realizing
that, in the end, in a user-driven project like \astropy, what happens
depends mostly on what is highest on the priority list of those that
have time to contribute).

\paragraph{Uncertainties and Distributions.} Early in our careers, we
were likely all taught that a measurement without a unit or an
uncertainty is essentially meaningless.  \astropyunits already
provides the ability to associate numbers with units, and propagate
these correctly, with its \astropyQuantity\ class, but so far the
ability to associate and propagate uncertainties is relatively
limited: in \astropysubpkg{nddata}, there are options to associate
errors with data arrays, but these are treated as independent, i.e.,
covariances are not tracked.

In general, proper error propagation is difficult, and often it is
best to use Monte Carlo methods.  With the new
\astropysubpkg{uncertainties} sub-package, a first step is made to
make this possible.  It allows one to generate, for each variable,
randomly drawn samples in a \astropyDistribution, and then propagate
these by passing them through the normal analysis, producing a
\astropyDistribution\ of final results that can be inspected.

This \astropyDistribution\ can be any type of array, including of
course a \astropyQuantity, and the goal is to ensure it can be
seamlessly used to instantiate also other \astropypkg classes, such as
\astropySkyCoord\ and \astropyTime.

Apart from Monte-Carlo error propagation, the plan for
\astropysubpkg{uncertainties} is to also support the case where values
have associated uncertainties, and these are propagated assuming that
they are normally distributed, but tracking covariances.  One
implementation problem to be solved is when to {\em stop} tracking
covariances.  For instance, a simple operation such as subtracting the
mean from $N$ data points implies that all data points are now
covariant with each other, i.e., one has to carry $N\times N$
covariances.  For a large image, that becomes not just pointless but
also prohibitive in terms of memory and CPU usage.

\paragraph{Masked quantities.}  Data can be known to be bad, in which
case it is useful to mask them.  Support for masked data involves a
number of choices, such as whether one indicates data values are bad
by a separate flag or by replacing them by a special value, such as
``Not a Number'' (NaN).  In \astropypkg, both approaches have been
used: a flag for masked columns in \astropyTable, mask flags and
bitmaps for N-dimensional data in \astropysubpkg{nddata}, and
replacing elements with NaN in \astropyTime.  A particular problem was
that no masked quantities were available, making it more difficult to
migrate to \astropyQTable, in which \astropyQuantity\ is used for all
columns with units, thus propagating those correctly (instead of
treating the unit merely as metadata).

Unfortunately, it turned out that the \texttt{MaskedArray} class from
\package{numpy} could not be used, as its design made it a wrapper
around any subclass, rather than allow another class to use it as an
alternative to a regular data array.  For quantities, this does not
work, as one can see from the fact that a masked value does not mean
the unit is not known.  Given this, a new \astropyMasked\ class was
designed, based on a framework very similar to that of
\astropyDistribution, making it easy to create masked instances of
other classes.  In particular, this can be used to create masked
quantities for \astropyQTable, including full support for reading and
writing to the different file types.

The masked quantities work mostly without any changes in higher-level
objects such as \astropySkyCoord, but work is still underway to expose
the mask in those classes, as well as to use masked arrays in
\astropyTime\ instead of replacing data with NaN.

\paragraph{Time series.}  From sampling a continuous variable at fixed
times to counting events binned into time windows, many different
areas of astrophysics require the manipulation of time series.  These
are often logically tables, and the new \astropysubpkg{timeseries}
sub-package extends the \astropyQTable\ class to tables of data as a
function of time, where the data can either represent samples or
averages over particular time bins.  The new classes offer a number of
special methods to manipulate time series (folding, resampling, etc.)
and to read different data formats (such as {\em Kepler} lightcurves,
etc.).

Also part of \astropysubpkg{timeseries} are common analysis routines,
including Lomb-Scargle and box-least-squares periodogrammes.

\paragraph{Spectral Coordinates.} Measurements are often taken at
specific ``spectral coordinates'', be they frequencies, wavelengths,
or photon energies.  The new \astropySpectralCoord\ class allows to
transform between these as well as to equivalent Doppler velocities.
It builds on the \astropyQuantity, which can also represent these
kinds of physical quantities, and allows conversion via dedicated
equivalencies, but aims to make this more straightforward.
Furthermore, it can be made aware of the observer and target reference
frames, allowing for instance transformation from telescope-centric (or
topocentric) frames to Barycentric or Local Standard of Rest
(LSRK and LSRD) velocity frames.

\paragraph{A High-Level Interface to World Coordinate Systems.}
Astronomical data are often provided with information about the
``real-world'' coordinates that correspond to pixel coordinates. This
mapping is the essence of the ``World Coordinate System (WCS)''
concept.  From its inception, the \astropysubpkg{wcs} sub-package
allowed access to WCS information provided in, e.g., FITS files, but
it became clear that other WCS standards and representations had to be
supported for new missions and observatories (e.g., the James Webb
Space Telescope and the Rubin Observatory).  To harmonize these, a new
high-level interface was created, based on a formal design first
proposed via an \astropy\ Proposal for Enhancement (\astropyAPE{14}).
The hope is that by having a formal design, it will be straightforward
for other packages implementing WCS objects to either modify their
classes to conform to the new interface or build thin wrappers that
conform to it.  The implementation in \astropysubpkg{wcs} interacts
well with other \astropypkg\ objects such as \astropySkyCoord\ and
\astropyTime.

\subsubsection*{Interoperability} \label{sec:core-features-interoperability}

  Because \astropy is modular and situated at the nexus between scientific
  computing and astronomy, interoperability is an important focus for \astropy,
  both within the various \astropypkg sub-packages and with other important
  \python libraries. If done right, all a user notices is that code using many
  different python features and libraries ``just works''. Here, we describe a
  few particular efforts towards this goal.

  \paragraph{\package{numpy} on Units}

    \astropyQuantity\ is a backbone of \astropy, leveraging the power of
    \package{numpy} and adding units. Previously, many \package{numpy} functions
    would strip a \astropyQuantity\ of its units (or fail outright), limiting
    \astropyQuantity's usefulness. Now, advances in \package{numpy} function
    overloading (see for example
    \href{https://numpy.org/neps/nep-0018-array-function-protocol.html}{NEP 18})
    mean \astropyQuantity\ works with almost all \package{numpy} (version 1.17+)
    functions.

    In the remaining gaps, the \package{numpy} and \astropyQuantity\
    interoperability efforts are ongoing. For some functions, such as in
    \href{https://numpy.org/doc/stable/user/basics.rec.html#module-numpy.lib.recfunctions}{\package{numpy}'s
    module} for manipulating structured arrays, compatibility only requires
    extending the existing \package{numpy}-\astropypkg bridge framework.
    Community interest, in the form of a Feature Request (or better yet Pull
    Request) would be sufficient to see this realized. For a few remaining
    functions, discussed further in the
    \href{https://docs.astropy.org/en/stable/known_issues.html#known-deficiencies}{\astropy
    documentation}, the \package{numpy} framework does not yet allow for full
    interoperability with \astropy. A goal of \astropy, and the scientific
    \python community is to enhance and implement the frameworks, allowing
    \astropyQuantity\ to propagate units seamlessly across the whole ecosystem
    of \package{numpy}-like projects.

  \paragraph{\astropypkg on Units}

    Units were also integrated further within \astropypkg. For already-defined
    \astropyModel, for instance imported from another library, \astropymodeling
    can now add units to otherwise unit-less models.
    % TODO! transition words
    A new module – \astropycosmologyunits – has been added to the cosmology
    sub-package for defining and collecting cosmological units and
    equivalencies. The unit \texttt{littleh} and equivalency
    \astropyapidoc{cosmology.units.with_H0}{with\_H0} are deprecated from the
    main \astropyunits sub-package and moved to \astropycosmologyunits. A new
    unit, \texttt{redshift}, is added for tracking factors of cosmological
    redshift. To correctly use redshift units in dimensional analysis,
    \texttt{redshift} is
    \astropyapidoc{cosmology.units.dimensionless_redshift}{treated as
    dimensionless} by default. To convert between redshift and other
    cosmological distance measures, e.g. CMB temperature or comoving distance,
    the equivalency
    \astropyapidoc{cosmology.units.with_redshift}{with\_redshift}
    is also added. This equivalency is actually a composite of other
    equivalencies, which may be used separately.

  \paragraph{Table Mixin Columns}

    % TODO: this is very rough. Probably needs a bit of a re-write.

    Within \astropypkg and \python there are numerous ways to represent and
    store array-valued data. Some of the differences are historical:
    \astropyapidoc{table.Column}{table.Column} and
    \astropyapidoc{io.fits.Column}{io.fits.Column} are not the same. Some
    differences are computational: \package{numpy}, \package{cupy}, and
    \package{dask} arrays are almost API identical, but are optimized for
    different compute needs. Lastly, some differences are inherent:
    \astropyQuantity, \astropyTime and \astropySkyCoord represent fundamentally
    different types of objects.

    \astropy believes however array-valued data is stored, it should be usable
    as a column in a table: historical differences are baked into stable API,
    computational needs can differ, and some objects have extra information. The
    \astropytable and \astropyfits table implementations allow for mixin-columns
    (in general a mixin column may contain multiple data components as well as
    object attributes beyond the standard column). A well defined protocol has
    been developed to allow the usage of these mixin columns while allowing the
    object to ``round-trip'' through tables with no loss of data or attributes.
    With this protocol, it is now possible to store not only standard
    \astropyTableColumn objects to a FITS table
    \href{https://docs.astropy.org/en/stable/io/fits/api/hdus.html}{HDU}, but
    also any \astropy native objects within a \astropyTable. This includes
    \astropyTime, \astropyQuantity, \astropySkyCoord, and many others. For
    objects not already covered by the mixin protocol, functions can be
    registered with \astropyTable to convert, on-the-fly, any array-like object
    into a `mixin' column that can be used in a table. As an example, this is
    used to provide the seamless integration of \package{dask} arrays in
    \astropyTable, allowing computer cluster-scale data to be stored in a
    \astropyTable.

  \paragraph{Astropy Fits in Time}

    The FITS standard was extended to rigorously describe time coordinates in
    the World Coordinate System (WCS) framework \citep{FITS-Time:2015}. Compared
    to other coordinates in the WCS framework, time requires more metadata --
    format, scale, position, reference etc. -- which had to be manually
    specified and understood. \astropyfits can store this data in a standard
    \astropyFitsColumn (for details, see the
    \href{https://docs.astropy.org/en/stable/io/fits/index.html}{FITS
    documentation}), but will not ``understand'' the metadata and attributes.
    Thankfully, through the support of the
    \href{https://summerofcode.withgoogle.com/archive/2017/projects/4778482366152704}{Google
    Summer of Code} 2017 program, the \astropyfits package now allows
    \astropyTime as a `mixin' column. \astropyTime columns in FITS tables
    ``understand'' the metadata while also providing a similar interface as the
    regular \astropyFitsColumn objects. With \astropyTime columns, \astropy
    ensures that time coordinates are stored according to the FITS standard in a
    consistent manner and can be read/written without loss of information. For
    backward compatibility with manual systems, this feature may be turned off,
    but it is recommended to update. To learn more about this project please see
    the final report,
    \href{https://aaryapatil.wordpress.com/2017/08/28/a-mixin-protocol-for-seamless-interoperability/}{A
    mixin protocol for seamless interoperability}.

  \paragraph{Persistent Storage and I/O}

    % TODO! need good transition sentence. This one is bad.
    The column mixin options in tables are mirrored in the increased support for
    I/O options.

    \begin{itemize}
      \item ECSV : Support for multidimensional and object columns in ECSV
      \item FITS : Support for writing Dask arrays to FITS files, and Append table to existing FITS file
      \item MRT Support for reading and writing tables to MRT format : GSOC
      \item Support for reading and writing tables to Parquet format
      \item Support for reading and writing tables to QDP format
      \item ASDF read write
    \end{itemize}

    Also stuff that isn't a table

    \begin{itemize}
      \item Support for reading, writing, and converting Cosmology
    \end{itemize}

  \paragraph{Unified I/O architecture}

    % TODO! better transition
    \astropyiosubpkg{registry} is a powerful way to define input and output
    (I/O) functions for \astropypkg objects, such as reading or writing to a
    file. \astropy uses this internally for the I/O methods in \astropyTable and
    \astropyCosmology (I/O added v5.0+), and users can register custom I/O to
    extend the options on these classes.

    As of Astropy 5.0 the I/O registry submodule has switched to a class-based
    architecture and been refactored to allow read-only, write-only, and
    read-write registries. Additionally, the registry submodule no longer
    assumes an object has \astropy-style \texttt{read} and \texttt{write}
    methods when trying to read or write that object. These two improvements
    enable a number of new use cases for \astropyiosubpkg{registry}, some
    highlighted here, but others we hope to see have community adoption.

    First, the class-based architecture allows for the creation of custom
    registries. An application of this is in \astropycosmology, which has two
    different \textit{kinds} of registries for \astropyCosmology: one for
    reading and writing files, and another for converting between python
    objects. Second, \textit{any} kind of object is now compatible with
    \astropyiosubpkg{registry}. Using \astropyiosubpkg{registry} for custom
    classes means a user does not have to create a whole new I/O system and can
    offer a unified and familiar API.

\subsection*{Precision, Accuracy, \& Reproducibility} \label{sec:core-features-precision_accuracy_repreducibility}

  An important focus of \astropy has been to increase the precision and accuracy
  of its functionality, while, where appropriate, making sure results obtained
  using older versions of \astropypkg are reproducible. Within each section we
  include in our descriptions how old results may be reproduced. The primary
  means are with configuration file options, and settings on runtime
  configuration objects called \astropyScienceState, described further in the
  documentation. Notable improvements to \astropypkg have been made to
  \astropytime, \astropyconstants, \astropycoordinates, and \astropycosmology,
  and we describe each in turn.

  \paragraph{Time}
    In astronomy, time accuracy down to the (nano)second is frequently
    important. Whether for planning observations, crunching pulsar or VLBI data,
    missing seconds meaningfully impact results. Before, when using \astropyTime
    in \astropytime leap seconds had to be manually applied; now leap seconds
    are applied automatically. Moreover, \astropypkg updates internal time
    data-files to ensure that the correct leap second adjustment is always used.
    For reproducibility, e.g. with an old code, \astropyLeapSeconds may also be
    manually applied.

  \paragraph{Constants}

    Measurements of physical constants (and the units defined from them) improve
    over time. Periodically the standardized systems of units and constants are
    updated to reflect these improvements, for instance in 2019 the SI system was
    redefined \citep{NIST2019}, with an accompanying update to the physical
    constants in
    \href{https://codata.org/initiatives/data-science-and-stewardship/fundamental-physical-constants/}{\texttt{SI/CODATA
    2018}}. \astropypkg contains two modules -- \astropyconstants and
    \astropyunits -- for working with constants and units, respectively.
    \astropyconstants now defaults to use the \texttt{SI/CODATA 2018} values, with
    the relation between units in \astropyunits based upon these constants. Most
    of \astropypkg and affiliate packages build upon \astropyunits, so this update
    has far-reaching effects across the \astropy ecosystem and astronomy
    community.

    For reproducibility, \astropypkg allows the constants (and therefore units)
    definitions to be rolled back to prior values, for instance the
    \texttt{SI/CODATA 2014} definitions.
    % CITATION
    The constants and units definitions are set by configuration file as well as
    \astropyScienceState (an runtime configuration controller). For work where
    updates to the values of the fundamental constants significantly alters
    results we recommend including with the work an \astropy configuration file,
    which specifies the set of constants used.

  \paragraph{Cosmology}
    % TODO! At least one good sentence about "what is the benefit of astropy.cosmology"

    \astropycosmology contains classes for representing cosmological models.
    Bundled with the classes are commonly used \astropyCosmology realizations, for
    example best-fit measurements from
    \href{https://lambda.gsfc.nasa.gov/product/map/current/}{WMAP}
    \citep{WMAP2003} and \href{https://www.nasa.gov/mission_pages/planck}{Planck}
    \citep{PlanckMission:2006}.

    When these important missions publish new measurements, \astropycosmology is
    updated to include these results as realizations of the appropriate class
    (generally \astropyFlatLambdaCDM). The models used by WMAP and Planck do not
    always exactly correspond to \astropycosmology classes, for instance the
    Planck 2018 results \citep{Planck2018VI} include massive neutrinos in
    $\Omega_{matter,0}$, while in \astropyFlatLambdaCDM this mass contribution is
    stored in a separate parameter. Consequently, while some parameter values
    appear different from the source paper, the \astropyCosmology realizations
    correctly reproduce the cosmological models.

    Since \citealt{astropy:2018} the following cosmology realizations have been
    added:
    \href{http://docs.astropy.org/en/stable/api/astropy.cosmology.WMAP1.html}{\texttt{WMAP1}}
    from WMAP First Year \citep[Table 7: WMAP + CBI + ACBAR + 2dFGRS +
    Lya]{WMAP1Year:2003};
    \href{http://docs.astropy.org/en/stable/api/astropy.cosmology.WMAP3.html}{\texttt{WMAP3}}
    from WMAP Three Year \citep[Table 6: WMAP + SNGold]{WMAP3Year:2007};
    \href{http://docs.astropy.org/en/stable/api/astropy.cosmology.Planck15.html}{\texttt{Planck15}}
    from Planck 2015 \citep[Table 4: TT, TE, EE + lowP + lensing +
    ext]{Planck2015XIII:2016};
    \href{http://docs.astropy.org/en/stable/api/astropy.cosmology.Planck18.html}{\texttt{Planck18}}
    from the Planck 2018 best-fit cosmological parameters \citep[Table 2: TT, TE,
    EE + lowE + lensing + BAO]{Planck2018VI:2020}.

    Astropy provides a
    \href{http://docs.astropy.org/en/stable/api/astropy.cosmology.default_cosmology.html}{configurable
    default cosmology}. All calculations done in a cosmological context will use
    this default cosmology. The default \astropyCosmology has been updated to the
    Planck 2018 parameters. For reproducibility, the default cosmology may be set
    to old defaults, such as the WMAP Three Year values. The default cosmology is
    dynamically configurable using \astropyScienceState, meaning a set of
    calculations may be run using different assumed cosmologies, and results
    compared between the two.
    % Repeat the word default less.

  \paragraph{Coordinates}

    Being able to describe the position of objects, i.e. their coordinates, is
    fundamental across many knowledge domains. \astropypkg provides
    \astropycoordinates for working with low-level positional and velocity data
    all the way to high-level coordinate objects with reference frames,
    atmospheric information, and more. A central feature of \astropycoordinates is
    the ability to transform data between reference frames, e.g. \astropyICRS
    \citep{ICRS:1997} to \astropyGalacticLSR \citep{GalacticLSR:2010}. However,
    there have been some historical limitations for spatially proximate
    transformations, impacting the usefulness of \astropycoordinates for
    ground-based telescopes and intra-solar astronomy. Transformations in the
    \astropyAltAz frame were reasonably precise for very distant objects, but
    wrong by up around several arcseconds for e.g. the location of the moon. Now
    these transformations are much more precise, down to the milliarcsecond
    level. Similar precision improvements were made to the
    \astropyapi{coordinates.builtin_frames.HADec}{Hour Angle-Declination} frame
    transformations. Additionally,
    \astropyapi{coordinates.builtin_frames.BaseEclipticFrame}{Ecliptic frames}
    and associated transformations have been updated to correctly reflect the
    “true” (includes precession but not nutation) and “mean” (precession and
    nutation) terminology.

    For Galactic astronomers, the \astropyGalactocentric
    \astropyapi{coordinates.galactocentric_frame_defaults}{frame defaults} have
    been updated to include more recent measurements (tabulated in the
    \href{https://docs.astropy.org/en/stable/api/astropy.coordinates.galactocentric_frame_defaults.html#astropy.coordinates.galactocentric_frame_defaults.references}{\texttt{frame\_defaults.references}}
    attribute). For reproducibility, old definitions are still available, and will
    continue to be so as \astropypkg is updated with forthcoming measurements of
    the Galactic center, solar motion, etc.
    % may be used by setting the default value to ``pre-v4.0'' in the
    % \href{https://docs.astropy.org/en/stable/api/astropy.coordinates.galactocentric_frame_defaults.html}{\texttt{galactocentric\_frame\_defaults}}
    % \astropyScienceState.

\subsection*{Performance} \label{sec:core-features-performance}

As features are added and corner cases dealt with, performance often
suffers.  Hence, pull requests that improve performance are welcomed
generally, and were even the main goal for one release cycle.  For
most packages, this meant improvements in the \python\ code, but for
some the most time-critical pieces were rewritten in~C (e.g., for
convolution of images, sigma-clipping, and converting time strings to
binary).  Furthermore, a particular effort was made to ensure
\astropypkg\ is thread-safe, so that it can be used on supercomputer
clusters.
